---
title: "Assignment 4: Sensitivity Analysis Using Latin Hypercube Sensitivity (LHS)"
description: |
  This assignment analyzes model sensitivity. 
author: "Pol Carbo-Mestre & Trevor Maggart"
output:
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
    collapsed: yes
    smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align='center')
library(sensitivity)
library(tidyverse)
library(lhs)
library(purrr)
library(here)
library(ggpubr)
```
  
  
## Creating Our Samples

Using the Latin Hypercube Sensitivity (LHS) method, we generate samples for our four parameters in our function: wind speed (m/s), vegetation height (m), roughness (Ko), and zero plane displacement (Kd). We then create an object `parm` to store the parameter values for analysis.

```{r}
# setting a random seed 
set.seed(1)
pnames = c("v", "height", "k_o", "k_d")

# set the paramter length
npar = length(pnames)

# select the number of samples
nsample = 100

# create matrix with LHS values based on sample size and parameter names
parm_quant = randomLHS(nsample, npar)
colnames(parm_quant)=pnames

# create a dataframe to store values
parm = as.data.frame(matrix(nrow=nrow(parm_quant), ncol=ncol(parm_quant)))
colnames(parm) = pnames

# fill in the columns in the df with the distributed values
parm[,"v"] = qnorm(parm_quant[,"v"], mean = 250, sd=30)
parm[,"height"] = qunif(parm_quant[,"height"], min = 9.5, max = 10.5)

# set the standard deviation for the k_o and k_d values 
pvar = 0.01
parm[, "k_o"] = qnorm(parm_quant[, "k_o"], mean = 0.1, sd = 0.1*pvar)
parm[, "k_d"] = qnorm(parm_quant[, "k_d"], mean = 0.7, sd = 0.7*pvar)
```
  
## Running our `Catm.R` model

With the LHS values, we can now run our model. From here, we `cbind()` conductance results with input parameters, so that we can visualize our data for sensitivity analysis.

```{r}
# read in the catm.r function
source(here("Catm-1.R"))

# use pmap to run the function over all values
results <- pmap(.l = parm, .f = Catm)

results_df <- as.data.frame(results) %>% 
  pivot_longer(everything()) %>% 
  select(value) %>% 
  rename(cond_value = value)

# cbind conductance results with input parameters
cond <- cbind.data.frame(parm, results_df)

# use `gather()` to pivot the parameters
cond_tidy = cond %>%  gather(v, height, k_o, k_d, value="parmvalue", key="parm")
```

## Plotting the results

```{r}
# Density plots to display parameter uncertainty input values
p1<-ggplot(data = cond, aes(x = height)) +
    theme_linedraw()+
  theme(
    axis.title.x = element_text(face = 'bold', size = 9.5, vjust = -0.5, family = "Avenir"),
    axis.title.y = element_text(face = 'bold', size = 9.5, vjust = 2, family = "Avenir"),
    axis.text = element_text(size = 7, family = "Avenir", face = "bold"),)+
  geom_density() + labs(x = "Vegetation Height (m)", y = "Density\n") 
p2<-ggplot(data=cond, aes(x=v)) +
    theme_linedraw()+
    theme(
      axis.title.x = element_text(face = 'bold', size = 9.5, vjust = -0.5, family = "Avenir"),
      axis.title.y = element_text(face = 'bold', size = 9.5, vjust = 2, family = "Avenir"),
      axis.text = element_text(size = 7, family = "Avenir", face = "bold"),)+
  geom_density() + labs(x = "Windspeed (m/s)", y = "Density\n") 
p3<-ggplot(data=cond, aes(x=k_o)) +
    theme_linedraw()+
    theme(
      axis.title.x = element_text(face = 'bold', size = 9.5, vjust = -0.5, family = "Avenir"),
      axis.title.y = element_text(face = 'bold', size = 9.5, vjust = 2, family = "Avenir"),
      axis.text = element_text(size = 7, family = "Avenir", face = "bold"),)+
  geom_density() + labs(x = "Ko", y = "Density\n") 
p4<-ggplot(data=cond, aes(x=k_d)) +
    theme_linedraw()+
    theme(
      axis.title.x = element_text(face = 'bold', size = 9.5, vjust = -0.5, family = "Avenir"),
      axis.title.y = element_text(face = 'bold', size = 9.5, vjust = 2, family = "Avenir"),
      axis.text = element_text(size = 7, family = "Avenir", face = "bold"),)+
  geom_density() + labs(x = "Kd", y = "Density\n") 
## Arranging our density plots
ggarrange(p1, p2, p3, p4)
```

```{r}
## Labels for facet wrap
parm_labs <- c("Height (m)", "Kd", "Ko", "Wind speed (m/s)")
names(parm_labs) <- c("height", "k_d", "k_o", "v")
## Plotting our parameter sensitivity
ggplot(data = cond_tidy, aes(x = parmvalue, y = cond_value))+
  geom_point(col = 'darkslategray4', alpha = 0.9, size = 1.5) +
  geom_smooth(method=lm, color = "darkslategray", size = 0.5, fill = NA) +
  geom_smooth(method=lm, color = NA, size = 0.5, fill = "darkslategray", alpha = 0.2) +
  facet_wrap(~parm, scales="free", ncol=4, 
             labeller = labeller(parm = parm_labs)) +
  labs(y = "Conductance (mm/s)",
       x = "Parameter Value") +
theme_linedraw()+
  theme(
    axis.title.x = element_text(face = 'bold', size = 9.5, vjust = -0.5, family = "Avenir"),
    axis.title.y = element_text(face = 'bold', size = 9.5, vjust = 2, family = "Avenir"),
    axis.text = element_text(size = 7, family = "Avenir", face = "bold"),
    panel.grid.minor.y = element_blank(),
    strip.text = element_text(face = 'bold', family = "Avenir"),
    strip.background = element_rect(fill = "darkslategray"),
    legend.position = "none"
  )
```

```{r}
sen_result <- pcc(X = parm, y = results_df$cond_value, rank = TRUE)

sen_result_plot <- sen_result[[7]][1] 
sen_result_plot$parameter <- as.factor(row.names(sen_result_plot))
sen_result_plot$parameter <- fct_reorder(sen_result_plot$parameter, -sen_result_plot$original)

# plot of results
ggplot(sen_result_plot, aes(x = parameter, y = original, label = round(original,3))) +
  theme_linedraw() +
  geom_point(size=4, alpha = 0.9, color = "darkslategray4") +
  geom_text(hjust=-0.3, vjust=0, color = "darkslategray", family = "Avenir") +
  labs(title = "Partial Rank Correlation Coefficient Values",
       x = NULL,
       y = "Values\n") +
  theme(plot.title = element_text(hjust = 0.5, family = "Avenir"),
        axis.title.y = element_text(family = "Avenir"),
        axis.text.y = element_text(family = "Avenir"),
        axis.text.x = element_text(family = "Avenir"))
```
